%\documentclass[11pt][fleqn]{report}
\documentclass[fleqn,10pt]{book}

\usepackage{bm,amsmath}
\usepackage{makeidx}
\usepackage{color}
\usepackage{natbib}
\usepackage{mfpic}
\usepackage{pxfonts}
\usepackage{lastpage}
\usepackage{graphicx}
\usepackage{verbatim} 
\usepackage{authblk}
\usepackage{tikz}
\usepackage[framemethod=tikz]{mdframed}

\usepackage{titlesec}
\titleformat*{\section}{\LARGE\bfseries}
\titleformat*{\subsection}{\Large\bfseries}
\titleformat*{\subsubsection}{\large\bfseries}
\titleformat*{\paragraph}{\large\bfseries}
\titleformat*{\subparagraph}{\large\bfseries}
\renewcommand{\familydefault}{\sfdefault}

\newcommand{\bftau}{\mbox{\boldmath $\tau$}}


\usepackage{minitoc} 
\renewcommand{\mtcfont}{\small\sffamily} 
\renewcommand{\mtcSfont}{\small\sffamily} 
\usepackage[Lenny]{fncychap}

\evensidemargin -.3in
\oddsidemargin  0.0in

\topmargin -0.50in
\textheight 9.0in
\renewcommand{\baselinestretch}{1.0}
\textwidth 6.5in

\usepackage{fancyhdr}
\pagestyle{fancy}
\fancyhf{}
\lhead{\leftmark}
\rhead{Section \thesection}
\lfoot{{\sc Column tests of CVMix/KPP}}
\cfoot{\today}
\rfoot{Page \thepage}
\renewcommand{\footrulewidth}{1pt}
\usepackage{kpfonts}
\fancyhead[LE,LO]{\textsc{\textbf{\nouppercase{\leftmark}}}}

\usepackage{titlesec,titletoc} 
\titleclass{\part}{top}
\titleformat{\part}
{\centering\normalfont\Huge\bfseries}{}{0pt}{}

\usepackage{hyperref}  % must be used last in the list of packages 
\hypersetup{colorlinks=true,linkcolor=blue,citecolor=blue,filecolor=blue,urlcolor=blue} 


\setcounter{tocdepth}{1}
\setcounter{minitocdepth}{3}
\setcounter{secnumdepth}{3}

\makeindex 
\citeindextrue

\title{\sc Column tests of CVMix/KPP}

\date{\today}

% \author[$\star$]{Stephen M. Grif\/f\/ies \thanks{Stephen.Griffies@noaa.gov}}
% \affil[$\star$]{NOAA/Geophysical Fluid Dynamics Laboratory, Princeton USA}

% \author[$\dagger$]{Michael Levy \thanks{MLevy@ucar.edu}}
% \affil[$\dagger$]{National Center for Atmospheric Research, Boulder USA}

% \author[$\star$]{Alistair J. Adcroft \thanks{Alistair.Adcroft@noaa.gov}}
% \author[$\dagger$]{Gokhan Danabasoglu \thanks{Gokhan@ucar.edu}}
% \author[$\star$]{Robert W. Hallberg \thanks{Robert.Hallberg@noaa.gov}}

% \author[$\triangle$]{Doug Jacobsen \thanks{Jacobsen.douglas@gmail.com}}
% \affil[$\triangle$]{Los Alamos National Laboratory, Los Alamos USA}

% \author[$\dagger$]{William Large \thanks{Wily@ucar.edu}}
% \author[$\triangle$]{Todd Ringler \thanks{Ringler@lanl.gov}}


\author[$\star$]{The CVMix Team}

\begin{document}

\pagenumbering{roman}
\maketitle 
\thispagestyle{empty}


\begin{center}
{\scshape \Large Column tests of CVMix/KPP} 
\end{center}

This document details tests with the KPP \citep{LargeKPP} boundary
layer parameterization in CVMix.  The ocean primitive equations are
time stepped in the ocean models MOM6, MPAS-ocean, and POP, with each
model having implemented CVMix/KPP.  The models are configured in
two-dimensional (x,y) periodic domains so that there are no lateral
gradients.  Furthermore, each uses the same grid spacing and time
step.  However, time stepping schemes differ, thus precluding bitwise
answer agreements.  Indeed, answer differences can be far more than
roundoff, illustrating the sensitivity of the KPP scheme to time
stepping details.  Physical and numerical aspects of CVMix/KPP are
provided in {\sc Theory and Numerics of the Community Ocean Vertical
  Mixing (CVMix) Project}.  Salient points from that document are
assumed here.

As there are no lateral gradients, evolution of velocity and tracer
are one-dimensional in the vertical.  We generally initialize velocity
at zero, so that it evolves only when there is a nonzero surface
stress.  If tracers have a uniform initial profile, their evolution is
initiated by buoyancy forcing via surface heating.  Once a nontrivial
vertical profile is established, either through boundary fluxes or the
initial profile, then evolution is impacted by the local portion of
KPP realized as downgradient diffusion.  Additionally, when surface
buoyancy forcing is negative, tracers are impacted by the non-local
portion of KPP.  The non-local operator redistributes the negative
surface buoyancy forcing throughout the boundary layer.


\dominitoc
\tableofcontents
\listoffigures

\pagenumbering{arabic}

\include{KPP_test_common}
\include{KPP_test_wind_only}
\include{KPP_test_wind_and_heating}
\include{KPP_test_cooling}


%%%%%%%%%%%%%%%%%%%%%      References     %%%%%%%%%%%%%%%%%%%%%%%%%%%
\newpage 

\bibliographystyle{elsarticle-harv.bst}
\addcontentsline{toc}{chapter}{\scshape Bibliography} 
\markboth{\scshape BIBLIOGRAPHY}{\scshape BIBLIOGRAPHY}
\gdef\rightmark{\scshape BIBLIOGRAPHY}
\bibliography{Griffies_refs}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%








\end{document}
